---
title: "iHMP_metadata_analysis_2"
output: html_document
---

### Setup

```{r setwd, include=FALSE}
#setwd("R/2020-IBD-Virome/")
```

```{r trunc-output, include=FALSE}
# save the built-in output hook
hook_output <- knitr::knit_hooks$get("output")

# set a new output hook to truncate text output
knitr::knit_hooks$set(output = function(x, options) {
  if (!is.null(n <- options$out.lines)) {
    x <- xfun::split_lines(x)
    if (length(x) > n) {
      # truncate the output
      x <- c(head(x, n), "....\n")
    }
    x <- paste(x, collapse = "\n")
  }
  hook_output(x, options)
})
```


Load libraries
```{r load-libs, message=FALSE, warning=FALSE}
library(readr)
library(dplyr)
library(stringr)
```

Load Data
```{r load-data, message=FALSE, warning=FALSE}
og_metadata <- readr::read_csv("hmp2_metadata.csv")
```

### Explore metadata
Examine data:
703 viromics samples
```{r exmn-data}
barplot(table(og_metadata$data_type))
table(og_metadata$data_type)
```

Examine colnames:
```{r exmn-data-dims, out.lines=8}
print(og_metadata)
```
```{r exmn-col-names, out.lines=25}
colnames(og_metadata)
```

### Isolating metadata of interest
The first goal will be to isolate only subject IDs with viral data

isolate viral data from IBD project

```{r exmn-virdata, out.lines=25}
vir_metadata <- og_metadata %>%
    filter(data_type=="viromics") %>%
    rename(research_project=`Research Project`) %>%
    rename(participant_id=`Participant ID`) %>%
    rename(external_id=`External ID`) %>%
    rename(age_at_diagnosis=`Age at diagnosis`) %>%
    filter(research_project=="ibdmdb")
print(vir_metadata)
```


```{r, include=FALSE, results=FALSE}
# String searches didnt help id specific antibiotic bc the phrase "antibiotic" is not present in the table, just the whole names of the antibiotic. but it did help id relevant columns

#str_detect(string = colnames(viral_metadata), pattern = "ntibiot")
colnames(vir_metadata)[str_detect(string = colnames(vir_metadata), pattern = "ntibiot")]

```

I want to look at columns "Antibiotics", "Has the subject used any antibiotics since the last visit?", "In the past 6 months, have you used antibiotics?", "Other Antibiotic:", and "Reason for stopping other antibiotic:"

first, some column renaming
```{r}
vir_metadata <- vir_metadata %>%
    rename(abx=`Antibiotics`) %>%
    rename(abx_since_last_visit=`Has the subject used any antibiotics since the last visit?`) %>%
    rename(abx_past_6_mo=`In the past 6 months, have you used antibiotics?`) %>%
    rename(abx_other=`Other Antibiotic:`) %>%
    rename(abx_other_reason_for_stopping=`Reason for stopping other antibiotic:`)

```

the Antibiotic columns (except abx) dont have data at all, likely stored under a different -omics line, so will find that later. 
```{r}
exp_metadata <- vir_metadata %>%
    filter(abx=="Yes") %>%
    select(abx, participant_id)
print (exp_metadata)

exp_metadata <- vir_metadata %>%
    filter(!is.na(abx_since_last_visit))
print (exp_metadata)

exp_metadata <- vir_metadata %>%
    filter(!is.na(abx_past_6_mo))
print (exp_metadata)

exp_metadata <- vir_metadata %>%
    filter(!is.na(abx_other))
print (exp_metadata)

exp_metadata <- vir_metadata %>%
    filter(!is.na(abx_other_reason_for_stopping))
print (exp_metadata)

```
This is the list of viromics samples relevant to the study
```{r}
viromics_samples_metadata <- vir_metadata %>%
    filter(abx=="Yes") %>%
    arrange(participant_id, week_num)

print(viromics_samples_metadata)
```

This is the list of participants relevant to the study, theres 34 lol
```{r}
participants <- vir_metadata%>%
    filter(abx=="Yes") %>%
    arrange(participant_id, week_num) %>%
    select(participant_id) %>%
    distinct(participant_id)
print(participants)
```


# So now lets get that sample numbers, in a list

```{r}
participants_ids_list <- participants[,1]
print(typeof(participants))
```



