---
title: "Metadata Summary"
output: 
  html_document: 
    keep_md: yes
date: '2022-03-16'
---

### Definitions 
MVX - viromics, Viral fraction
MGX - metagenomic, Bulk fraction


### Chunk 1 - Load Libraries and Data 
```{r Load-Libraries-and-Data, warning=FALSE}
library(tidyverse)
library(janitor)
library(naniar)

#setwd("/Users/hehouts/projects/dynamic-duos/R")

ibdmdb <- clean_names(readr::read_csv("csvs_and_other_metadata/ihmp_metadata.csv", show_col_types = F))
ibdmdb <- ibdmdb %>% 
  filter(data_type == "metagenomics" | data_type == "viromics") %>% 
  arrange(participant_id, data_type, week_num)
```


### Chunk 2 - Determine Number of Metagenome and Viromes Samples
Conclusions:
* 2341 Metagenome & Virome samples from the iHMP IBD Study
  + 1638 Metagenomes 
  + 703 Viromes. 


```{r Determine-Number-of-Metagenome-and-Viromes-Samples}

# 2341 total samples, mgx + mvx
ibdmdb <- ibdmdb %>% 
  filter(data_type == "metagenomics" | data_type == "viromics") %>% 
  relocate(external_id, tube_a_viromics, week_num, participant_id, data_type) %>% 
  arrange(participant_id, data_type, week_num)

# 1638 MGX samples
ibdmdb_mgx_only <- ibdmdb %>% 
  filter(data_type == "metagenomics") %>% 
  arrange(participant_id, week_num)

# 703 MVX samples
ibdmdb_mvx_only <- ibdmdb %>% 
  filter(data_type == "viromics") %>% 
  arrange(participant_id, week_num)

# hard saved csv of all mgx & vrx samples in the ibdmdb, no other 'omics. should be 2341 long. 
write_csv(ibdmdb, "csvs_and_other_metadata/ibdmdb_metadata_2341.csv")
```

### Chunk 3 - Determine Number of participants with both MVX and MGX Samples, and Determine Which Have Antibiotic Data

Conclusions:
* 130 total participants in the study
* 105 participants have both virome and metagenome samples
  + 2299 samples are from those 105 participants
* 34 participants that have viromes samples, and had antibiotics
  + 838 samples are from those 34 participants

```{r}

# 130 total participants
ibdmdb %>% distinct(participant_id)

# 105 participant ids of ONLY participants that have vrx samples, regardless of antibiotic status, i.e. "pair" participants
pair_ids_vec <- ibdmdb %>% 
  select(data_type, participant_id) %>% 
  filter(data_type == "viromics") %>% 
  distinct(participant_id) %>% 
  pull(participant_id)

# 2299 samples from just the 105 "pair" participants
pairs_df <- ibdmdb %>% 
  filter(data_type == "metagenomics" | data_type == "viromics") %>% 
  filter(participant_id %in% pair_ids_vec) %>%
  arrange(participant_id, data_type, week_num)
#write_csv(pairs_df, "csvs_and_other_metadata/ihmp_metadata_2299_paired_mgx_vrx.csv")


### Antibiotics:

# 34 participant ids of ONLY participants that have vrx samples, and had antibiotics, i.e. "abxpair", "relevant participants"
abx_pair_ids_vec <- ibdmdb %>% 
  select(data_type, participant_id, antibiotics) %>% 
  filter(data_type == "viromics" & antibiotics=="Yes") %>%
  distinct(participant_id) %>% 
  pull(participant_id)


# 838 samples from just the 34 "abxpair" participants,
abx_pairs_df <- ibdmdb %>% 
  filter(data_type == "metagenomics" | data_type == "viromics") %>% 
  filter(participant_id %in% abx_pair_ids_vec) %>%
  arrange(participant_id, data_type, week_num)
#write_csv(abx_pairs_df, "csvs_and_other_metadata/ihmp_metadata_838_antibiotics_paired_mgx_vrx.csv")

```



### Chunk 4 - Finding existing filenames for viromes in the ibdmdb metadata

After some sleuthing, it was determined that the existing virome filenames, which start with the prefix `SM-`, correspond with the `tube_a_viromics` column
```{r}
ibdmdb %>% 
  select(contains("tube"))
```





### Chunk 5 - s5 subset: File IDs for viromes
I want to have the viromes and metagenomes for the following 5 participants. (2 have abx, 3 dont)

```
C3001
C3002
C3003
C3004
C3005
```

```{r}
ibdmdb %>% 
  filter(participant_id %in% c("C3001", "C3002", "C3003", "C3004", "C3005")) %>%
  mutate(abx_group = case_when(
      participant_id %in% abx_pair_ids_vec~T,
      !(participant_id %in% abx_pair_ids_vec)~F
      )) %>%
  select(participant_id, abx_group) %>% 
  unique()


subset5 <- ibdmdb %>% 
  filter(participant_id %in% c("C3001", "C3002", "C3003", "C3004", "C3005")) %>%
  mutate(abx_group = case_when(
      participant_id %in% abx_pair_ids_vec~T,
      !(participant_id %in% abx_pair_ids_vec)~F
      )) %>% 
  relocate(external_id, tube_a_viromics, week_num, participant_id, abx_group, data_type) %>% 
  select(1:15)
subset5

subset5_virome_filenames <- subset5 %>% 
  filter(data_type == "viromics")
```


### Chunk 6 - uids - a new ibdmdb table with download filenames and uid fro each sample

```{r}
ibdmdb_uids <- ibdmdb %>% 
  mutate(abx_group = case_when(
      participant_id %in% abx_pair_ids_vec~T,
      !(participant_id %in% abx_pair_ids_vec)~F
      )) %>% 
  relocate(external_id, tube_a_viromics, week_num, participant_id, abx_group, data_type
      ) %>% 
  select(1:12) %>% 
  mutate(file_prefix = case_when(
    (data_type == "viromics")~tube_a_viromics,
    (data_type == "metagenomics")~external_id
    )
  ) %>% 
  mutate(zip_type = case_when(
        (data_type == "viromics")~"VIR",
        (data_type == "metagenomics" &
           grepl('_P$', external_id) == T)~"P",
        (data_type == "metagenomics" &
           grepl('_TR$', external_id) == T)~"TR",
        (data_type == "metagenomics" &
           !grepl('_TR$', external_id) &
           !grepl('_P$', external_id) == T)~"STD"
    )) %>% 
mutate(uid = case_when(
    (data_type == "metagenomics")~external_id,
    (data_type == "viromics")~tube_a_viromics,
    )) %>% 
  mutate(file_suffix = case_when(
    (zip_type == "VIR")~".tar",
    (zip_type == "STD")~".tar",
    (zip_type == "P")~".fastq.gz",
    (zip_type == "TR" )~".tar",
    )) %>%
  relocate(file_prefix, file_suffix) %>% 
  unite("file", file_prefix:file_suffix, sep = "", remove = F) %>% 
  relocate(file, zip_type, uid)
ibdmdb_uids

ibdmdb_uids_s5 <- ibdmdb_uids %>% filter(participant_id %in% c("C3001", "C3002", "C3003", "C3004", "C3005"))

ibdmdb_uids %>% select(file, file_prefix, zip_type, uid) %>% 
  write_csv("csvs_and_other_metadata/file_prefix_zip_uid.csv")

ibdmdb_uids %>% 
  filter(participant_id %in% c("C3001", "C3002", "C3003", "C3004", "C3005")) %>%
  select(file, file_prefix, zip_type, uid) %>%
  write_csv("csvs_and_other_metadata/file_prefix_zip_uid_s5.csv")


```


### Chunk 7 make the filename.txt's for all the zip subtypes

```{r}
# VIR
ibdmdb_uids %>% filter(zip_type == "VIR") %>% select(file) %>%  
  write_csv("../snakecrate/data/raw_data/vir/vir_filenames.txt", col_names = F)

ibdmdb_uids_s5  %>% filter(zip_type == "VIR") %>% select(file) %>%  
  write_csv("../snakecrate/data/raw_data/vir/s5_vir_filenames.txt", col_names = F)
######


# TR
ibdmdb_uids %>% filter(zip_type == "TR") %>% select(file) %>%  
  write_csv("../snakecrate/data/raw_data/tr/tr_filenames.txt", col_names = F)

ibdmdb_uids_s5  %>% filter(zip_type == "TR") %>% select(file) %>%  
  write_csv("../snakecrate/data/raw_data/tr/s5_tr_filenames.txt", col_names = F)
######


# P 
ibdmdb_uids %>% filter(zip_type == "P") %>% select(file) %>%  
  write_csv("../snakecrate/data/raw_data/p/p_filenames.txt", col_names = F)

ibdmdb_uids_s5  %>% filter(zip_type == "P") %>% select(file) %>%  
  write_csv("../snakecrate/data/raw_data/p/s5_p_filenames.txt", col_names = F)
######


# STD
ibdmdb_uids %>% filter(zip_type == "STD") %>% select(file) %>%  
  write_csv("../snakecrate/data/raw_data/std/std_filenames.txt", col_names = F)

ibdmdb_uids_s5  %>% filter(zip_type == "STD") %>% select(file) %>%  
  write_csv("../snakecrate/data/raw_data/std/s5_std_filenames.txt", col_names = F)
######


```



##Supplementary Chunks
### Chunk s1
```{r}
#### Here I am frantically looking for the column that its matching on 76C9Y, with external_id HSM6XRR7

# reducing the number of columns to char columns, gets down to ~300 from ~500 
df <- ibdmdb_mvx_only %>% 
  filter(external_id == "HSM6XRR7") %>% 
  select(where(is.character))


df2 <- df[!map_lgl(df, ~ all(is.na(.)))]
df2

# df2 only has 131 columns, so I just clicked through them. found:

#`tube_a_viromics`!!!! 
#which matches the exact virome file name: SM-76C9Y


# looking at these "tube" columns is proabably going to be important later, hard listed below 
#names(
  ibdmdb %>% 
  select(contains("tube")) %>% 
  select(where(is.character)) %>% 
  drop_na()
#)
```
```
"serum_tube_number_1_received_at_csmc"
"serum_tubes_number_2_4_received_at_mgh"
"tube_a_dna_rna""tube_a_metabolomics"
"tube_a_storage"
"tube_a_viromics"
"tube_b_fecal_calprotectin"
"tube_b_proteomics"
"stool_sample_id_tube_a_et_oh"
"sample_id_tube_b_no_preservative"
"tube_a_and_b_received_at_broad"    
```



### Chunk s2, Download some viromes:
```{r, eval=F}
subset5 %>% 
  mutate(
    download_URL = case_when(
      (data_type=="viromics")~"https://ibdmdb.org/tunnel/static/HMP2/Viromics/1732/"
      )
    ) %>% relocate(download_URL)
```





### Chunk s3, check that all in ibdmdb are unique external_ids
```{r}
#1638
ibdmdb %>% 
  filter(data_type=="metagenomics") %>% 
  distinct(external_id)

#703
ibdmdb %>% 
  filter(data_type=="viromics") %>% 
  distinct(external_id)

#2341
ibdmdb

#1656
ibdmdb %>% 
  distinct(external_id)

1638-703
#935

2341-935
#1406
```

### bash sandbox

```{bash}
while IFS="," read -r file prefix zip_type uid
do
  echo "file is $file"
  echo "file prefix is $prefix"
  echo "zip_type is $zip_type"
  echo "uid is $uid"
done < <(tail -n +2 csvs_and_other_metadata/file_prefix_zip_uid.csv)
```



```{bash}
while IFS="," read -r file prefix zip_type uid
do
echo "on file $file with type $zip_type .... "
  case $zip_type in
    "VIR")
      echo "virus"
      ;;
    "P"|"TR"|"STD")
      echo "mtg"
      ;;
  esac
done < <(tail -n +2 csvs_and_other_metadata/file_zip_uid.csv)
```



## SRA exploration
### Chunk SRA
Read in SRA results:

```{r}
sra <- clean_names(readr::read_csv("csvs_and_other_metadata/sra_result2979.csv"))
sra <- sra %>% 
  filter(organism_name == "human gut metagenome" | organism_name == "viral metagenome") %>% 
  filter(library_strategy == "WGS") %>% 
  relocate(sample_accession, library_name, organism_name, experiment_title) %>% 
  arrange(desc(organism_name))

ihmp_sras <- sra %>% select(sample_accession)
write_csv(ihmp_sras, "csvs_and_other_metadata/sras_mgx_vrx_2041.csv")

head(sra)
sra_cols<- names(sra)
```







does `all_samples$external_id` merge on a trimmed down `sra$library_name`!?!?!

```{r eval=FALSE}


#sum_all_samples <- all_samples %>% select(project, participant_id, week_num, data_type, site_sub_coll, external_id)



sra_mod <- sra %>% 
  separate(library_name, into = c("library_name", NA), sep = "_MGX", remove = T) %>% 
  separate(library_name, into = c("library_name", NA), sep = "_MVX", remove = T)
sra_mod

sra_mgx <- sra %>% 
  filter(organism_name == "human gut metagenome")
  
    


### use a split sra$experiment_title and ihmp$external_id to 
sra_test <- sra_mgx %>% 
  separate(experiment_title, into = c("key", NA), sep = " stool")

#ihmp_mod <- all_samples %>% 
#  select(project, participant_id, week_num, data_type, site_sub_coll, external_id) %>% 
#  separate(project, into = c("project", NA), sep = "_MGX", remove = T)%>% 
#  separate(project, into = c("project", NA), sep = "_MVX", remove = T)

ihmp_mod <- abx_pairs_df %>% 
  select(project, participant_id, week_num, data_type, site_sub_coll, external_id) %>% 
  separate(project, into = c("project", NA), sep = "_MGX", remove = T)%>% 
  separate(project, into = c("project", NA), sep = "_MVX", remove = T)




ihmp_mod_2 <- ihmp_mod %>% 
  separate(external_id, into = c(external_id_trim,""), sep = "",remove = T)



  
left_join(ihmp_mod, sra_mod, by = c("project" = "library_name")) %>%
  drop_na(sample_accession) %>% 
  relocate(sample_accession, project, organism_name, experiment_title)
```


### Chuck more SRA

does `all_samples$project` merge on `sra$library_name`!?!?! only on viromes

```{r eval=FALSE}


sum_all_samples <- all_samples %>% select(project, participant_id, week_num, data_type, site_sub_coll, external_id)

left_join(sum_all_samples, sra, by = c("project" = "library_name")) %>%
  drop_na(sample_accession) %>% 
  relocate(sample_accession, project, organism_name, experiment_title)
#%>% 
#  pull()



```



### end
